---
title: "pym_test"
author: "Dan Gunn"
date: "2023-08-02"
output: html_document
---

# The Narrative of Arthur Gordon Pym
This is a notebook to learn the core NLP package from Stanford and NLP generally.
Using the only novel by Edgar Allen Poe as an example. 
https://www.gutenberg.org/ebooks/51060
Importing Arthur Gordon Pym of Nantucket text from Project Gutenberg

```{r}
#installing libraries
library(tidyverse)
library(coreNLP)
library(cleanNLP)

#loading the English model
initCoreNLP("../NLP_library/stanford_core_models", type = c("english"), 
            parameterFile = NULL, 
            mem = "4g")
```
Importing Arthur Gordon Pym of Nantucket text from Project Gutenberg
```{r}
pym_full_raw <- read_delim("https://www.gutenberg.org/cache/epub/51060/pg51060.txt", delim = "\t")
```
Turning it into a character vector.
```{r}
pym_full_flat <- str_flatten(deframe(pym_full_raw))
```
Just choosing a small excerpt to work with.
```{r}
pym_excerpt <- "Augustus Barnard thoroughly entered into my state of mind. It is probable, indeed, that our intimate communion, Augustus and mine, had resulted in a partial interchange of character. About eighteen months after the period of the Ariel's disaster, the firm of Lloyd and Vredenburgh (a house connected in some manner with the Messieurs Enderby, I believe, of Liverpool) were engaged in repairing and fitting out the brig Grampus of Philadelphia for a whaling voyage."
```
#Creating an annotation object from the excerpt character vector.
```{r}
pym_ex_annotated <- annotateString(pym_excerpt)
str(pym_ex_annotated)

```
#core references show the sentence character starts /end etc.
```{r, R.options=list(max.print=10)}
getCoreference(pym_ex_annotated)
```
#shows the parsed values
```{r, R.options=list(max.print=10)}
getParse(pym_ex_annotated)
```
#gives dependency tags by location and governors.
```{r, R.options=list(max.print=10)}
getDependency(pym_ex_annotated)
```
#gives sentiment level data by sentence level!
```{r, R.options=list(max.print=10)}
getSentiment(pym_ex_annotated)
```
#getting tokens? 
```{r, R.options=list(max.print=10)}
getToken(pym_ex_annotated)
```
## Finding the persons
How to I access the tokens that have and entity tagged? 
```{r}
pym_ex_annotated$token$NER
```
```{r}
pym_tokens <- getToken(pym_ex_annotated)
pym_tokens %>%
  filter(NER != "O") #oddly Oh and not zero.
```
### Evaluating the NER output.
* It missed that Ariel was a vessel. 
* Lloyd and Vrendenburgh were both names and ORGs. 
* And it just didn't flag Grampus.


## Finding the Vessels


## Mapping the Sentiment
